---
title: "Benchmarks"
output:
  rmarkdown::html_vignette:
    code_folding: hide
params:
  run_benchmarks: false
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Benchmarks}
---

<style>
    img {
        border: 0;
        padding: 0;
        margin: 0;
    }
</style>


```{r}
#| include: false
knitr::knit_hooks$set(optipng = knitr::hook_optipng)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  optipng = "-quiet"
)
run_benchmarks <- params$run_benchmarks
```

The following plots compare the performance of [`moocore`][moocore] against [`emoa`][emoa] and [`bbotk`][bbotk]. Other R packages are not included in the comparison because they are based on these packages for the functionality benchmarked, so they are **at least as slow** as them. For example [`GPareto`][GPareto], [`mlr3mbo`][mlr3mbo], [`rmoo`][rmoo] and [`bbotk`][bbotk] use [`emoa`][emoa] to compute the hypervolume.  Not all packages provide the same functionality.

<details><summary>Show benchmarking setup code</summary>

```{r setup}
#| class.source: 'fold-show'
library(matrixStats)
library(data.table)
library(ggplot2)
library(moocore)

geomspace <- function(start, stop, num) {
  x <- round(10^(seq(log10(start), log10(stop), length.out = num)), 0)
  x[1] <- start
  x[length(x)] <- stop
  x
}

read_ndset <- function(filename, filter=FALSE) {
  cat("Get file '", filename, "'\n")
  destfile <- system.file(file.path("extdata", filename), package="moocore")
  if (destfile == "") {
    destfile <- file.path("../../../testsuite/data", filename)
    if (!file.exists(destfile)) {
      fileext <- if (endsWith(destfile, ".xz")) ".xz" else ""
      destfile <- withr::local_tempfile(fileext = fileext)
      base_url <- "https://github.com/multi-objective/testsuite/raw/refs/heads/main/data/"
      utils::download.file(paste0(base_url, filename), destfile, quiet = FALSE)
    }
  }
  x <- read_datasets(destfile)
  x <- x[, -ncol(x)] # Union of datasets
  if (filter)
    x <- filter_dominated(x)
  x
}

# This is adapted from atime:::plot.atime
benchmark_plot <- function (x, title = "", only_seconds=TRUE, ...) {
  expr.name <- N <- kilobytes <- NULL
  meas <- x[["measurements"]]
  by.dt <- meas[, x$by.vec, with = FALSE]
  tall.list <- list()
  for (unit.i in seq_along(x$unit.col.vec)) {
    col.name <- x$unit.col.vec[[unit.i]]
    unit <- names(x$unit.col.vec)[[unit.i]]
    if (is.null(unit) || unit == "")
      unit <- col.name
    tall.list[[unit.i]] <- meas[, data.table(N, by.dt,
      unit, median = get(col.name))]
  }
  tall <- rbindlist(tall.list)
  if (only_seconds) {
    tall <- tall[unit=="seconds", ]
    ylab <- "CPU time (seconds)"
  } else {
    ylab <- "median line, min/max band"
  }

  gg <- ggplot() + theme_bw(base_size=12) +
    geom_ribbon(aes(N, ymin = min, ymax = max, fill = expr.name),
      data = data.table(meas, unit = "seconds"), alpha = 0.25, show.legend=FALSE) +
    geom_line(aes(N, median, color = expr.name), data = tall) +
    geom_point(aes(N, median, color = expr.name), data = tall) +
    scale_x_log10(breaks = unique(tall$N), labels = unique(tall$N)) +
    scale_y_log10(ylab, labels = scales::label_log(), guide="axis_logticks") +
    labs(subtitle = title) +
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
  if (only_seconds) {
    gg <- gg + theme(legend.title = element_blank(),
      legend.background = element_rect(fill=alpha("white", alpha=0.5)),
      legend.key = element_rect(fill = NA, color = NA),
      legend.position = "inside",
      legend.position.inside = c(0.99, 0.01),
      legend.justification = c("right", "bottom"),
      legend.margin = margin_auto(0))
  } else {
    gg <- gg + theme(legend.title = element_blank(),
      legend.background = element_rect(fill=alpha("white", alpha=0.5)),
      legend.position = c(0.8, 0.625)) + facet_grid(unit ~ ., scales = "free")
  }
  gg
}

get_package_version <- function(package)
  paste0(package, " (", as.character(packageVersion(package)), ")")

benchmark <- function(name, x, N, setup, expr.list, prefix, title) {
  rds_file <- paste0("bench/bench-", prefix, "-", name, ".rds")
  if (run_benchmarks || !file.exists(rds_file)) {
    lapply(names(expr.list), library, character.only = TRUE)
    names(expr.list) <- sapply(names(expr.list), get_package_version, USE.NAMES=FALSE)
    res <- substitute(atime::atime(
      N = N,
      expr.list = expr.list,
      setup = SETUP,
      result=FALSE,
      times=5,
      seconds.limit=10), list(SETUP=setup))
    res <- eval(res)
    saveRDS(res, file = rds_file)
  } else {
    res <- readRDS(rds_file)
  }
  gg <- benchmark_plot(res, title = paste0(title, " for ", name))
  gg
}

get_ndset <- function(x, filter)
{
  if (is.null(x$generate)) return(read_ndset(x$file, filter=filter))
  do.call(moocore::generate_ndset, x$generate)
}

benchmark_all <- function(files, prefix, title, setup, expr.list, filter)
{
  for (name in names(files)) {
    p <- benchmark(name = name, x = get_ndset(files[[name]], filter=filter),
      N = files[[name]]$N, prefix = prefix, title = title,
      setup = setup,  expr.list = expr.list)
    print(p)
  }
}
```
</details>

# Identifying (non)dominated points

The following plots compare the speed of finding (non)dominated solutions,
equivalent to `moocore::is_nondominated()`, in 2D, 3D, 4D, 5D and 10D.  The
plots show that [`moocore`][moocore] is always 10 times faster than [`GPGame`][GPGame],
[`targeted`][targeted] and [`bbotk`][bbotk] for any number of objectives.  In addition,
[`GPGame`][GPGame] calculates wrong values with repeated coordinates
([vpicheny/GPGame#2](https://github.com/vpicheny/GPGame/issues/2)). The `rand4d` testcase is somewhat different from the
others, as it is mostly composed of dominated points (only 10590 points are
nondominated), while the other testcases are mostly nondominated sets.

```{r ndom}
#| fig.asp: 0.8
#| fig.width: 4.5
#| dpi: 150
#| fig.show: 'hold'
#| out.width :  '50%'
#| fig.align: 'default'
setup <- quote({
  stopifnot(nrow(x) >= N)
  z <- as.matrix(x[1L:N, ])
  tz <- t(z)
})

expr.list <- list(
  moocore = quote(which(moocore::is_nondominated(z))),
  bbotk = quote(which(bbotk::is_dominated(tz))),
  GPGame = quote(GPGame::nonDom(z)),
  targeted = quote(targeted::nondom(z))
)

files <- list(
  "test2D-200k"=list(file="test2D-200k.inp.xz", N=geomspace(1000, 50000, 10)),
  "ran3d-40k"=list(file="ran.40000pts.3d.1.xz", N=geomspace(1000, 25000, 10)),
  "ran4d"=list(file="ran.9000pts.4d.10.xz", N=geomspace(1000, 60000, 10)),
  "convex-4d"=list(generate=list(25000, 4, "convex-sphere", 42), N=geomspace(500, 25000, 10)),
  "sphere-5d"=list(generate=list(20000, 5, "sphere", 42), N=geomspace(100, 20000, 10)),
  "rmnk-10d"=list(file="rmnk_0.0_10_16_1_0_ref.txt.xz", N=geomspace(100, 20000, 10))
)

benchmark_all(files, prefix="ndom", title = "is_(non)dominated()",
  setup = setup, expr.list = expr.list, filter=FALSE)
```


# Exact computation of hypervolume

The following plots compare the speed of computing the hypervolume indicator in 3D, 4D, 5D and 6D.

```{r hv}
#| fig.asp: 0.8
#| fig.width: 4.5
#| dpi: 150
#| fig.show: 'hold'
#| out.width: '50%'
#| fig.align: 'default'
setup <- quote({
  ref <- colMaxs(x, useNames = FALSE) + 1L
  stopifnot(nrow(x) >= N)
  z <- x[1L:N, ]
  tz <- t(z)
})

expr.list <- list(
  moocore = quote(moocore::hypervolume(z, ref = ref)),
  emoa = quote(emoa::dominated_hypervolume(tz, ref = ref)))

files <- list(
  "DTLZLinearShape.3d"=list(
    file = "DTLZSphereShape.3d.front.1000pts.10",
    N = geomspace(200, 6500, 10)),
  "DTLZLinearShape.4d"=list(
    file = "DTLZLinearShape.4d.front.1000pts.10",
    N = geomspace(100, 2000, 10)),
  "DTLZLinearShape.5d"=list(
    file = "DTLZLinearShape.5d.front.500pts.10",
    N = geomspace(100, 1000, 10)),
  "DTLZLinearShape.6d"=list(
    file = "DTLZLinearShape.6d.front.700pts.10.xz",
    N = geomspace(100, 1000, 10))
)

benchmark_all(files, prefix="hv", title = "HV Computation",
  setup = setup, expr.list = expr.list, filter=TRUE)
```

As the plots show, [`moocore`][moocore] is always faster than [`emoa`][emoa] and, hence, faster than [`GPareto`][GPareto], [`mlr3mbo`][mlr3mbo], [`rmoo`][rmoo] and [`bbotk`][bbotk].

# Hypervolume contribution

The only R package, other than [`moocore`][moocore], able to compute hypervolume contributions `moocore::hv_contributions()` is [`emoa`][emoa]. However, [`emoa`][emoa] is buggy and calculates wrong values ([olafmersmann/emoa#1](https://github.com/olafmersmann/emoa/issues/1)).

[moocore]: https://multi-objective.github.io/moocore/r/
[GPGame]: https://cran.r-project.org/package=GPGame
[GPareto]: http://cran.r-project.org/package=GPareto
[bbotk]:  https://cran.r-project.org/package=bbotk
[emoa]: https://cran.r-project.org/package=emoa
[mlr3mbo]: https://cran.r-project.org/package=mlr3mbo
[rmoo]: https://cran.r-project.org/package=rmoo
[targeted]: https://cran.r-project.org/package=targeted
