---
title: "Computing Multi-Objective Quality Metrics"
bibliography: REFERENCES.bib
link-citations: true
---

<style>
.table {
  width: fit-content !important;
}
</style>

```{r}
#| include: false
knitr::knit_hooks$set(optipng = knitr::hook_optipng)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  optipng = "-quiet"
)
```

Several examples of computing multi-objective unary quality metrics.

```{r setup}
library(moocore)
library(dplyr, warn.conflicts=FALSE)
get_dataset <- function(filename)
  read_datasets(system.file(file.path("extdata", filename),
    package = "moocore", mustWork = TRUE))
apply_within_sets <- function(x, sets, FUN, ...) {
  FUN <- match.fun(FUN)
  sapply(X = split.data.frame(x, sets), FUN = FUN, ...)
}
```

# Comparing two multi-objective datasets using unary quality metrics

First, read the datasets:

```{r}
spherical <- get_dataset("spherical-250-10-3d.txt")
uniform <- get_dataset("uniform-250-10-3d.txt")

spherical_objs <- spherical[, -ncol(spherical)]
spherical_sets <- spherical[, ncol(spherical)]
uniform_objs <- uniform[, -ncol(uniform)] / 10
uniform_sets <- uniform[, ncol(uniform)]
```

Create reference set and reference point:

```{r}
ref_set <- filter_dominated(rbind(spherical_objs, uniform_objs))
ref_point <- 1.1
```

Calculate metrics:

```{r, results='asis'}
uniform_igd_plus <- apply_within_sets(uniform_objs, uniform_sets,
  igd_plus, ref = ref_set)
spherical_igd_plus <- apply_within_sets(spherical_objs, spherical_sets,
  igd_plus, ref = ref_set)

uniform_epsilon <- apply_within_sets(uniform_objs, uniform_sets,
  epsilon_mult, ref = ref_set)
spherical_epsilon <- apply_within_sets(spherical_objs, spherical_sets,
  epsilon_mult, ref = ref_set)

uniform_hypervolume <- apply_within_sets(uniform_objs, uniform_sets,
  hypervolume, ref = ref_point)
spherical_hypervolume <- apply_within_sets(spherical_objs, spherical_sets,
  hypervolume, ref = ref_point)

knitr::kable(data.frame(
  Uniform = c(mean(uniform_hypervolume), mean(uniform_igd_plus),
    mean(uniform_epsilon)),
  Spherical = c(mean(spherical_hypervolume), mean(spherical_igd_plus),
    mean(spherical_epsilon)),
  row.names = c("Mean HV","Mean IGD+","Mean eps*")), format = "html") %>%
  kableExtra::kable_styling(position = "center",
    bootstrap_options = c("striped", "condensed", "responsive"))
```

<br>

# IGD and Average Hausdorff are not Pareto-compliant

Example 4 by @IshMasTanNoj2015igd shows a case where IGD gives the wrong answer:

```{r example4_igd_plus}
#| fig.asp: 0.8
#| fig.width: 4.5
#| dpi: 150
#| fig.align: 'default'
ref <- matrix(c(10, 0, 6, 1, 2, 2, 1, 6, 0, 10), ncol = 2, byrow=TRUE)
A <- matrix(c(4, 2, 3, 3, 2, 4), ncol = 2, byrow=TRUE)
B <- matrix(c(8, 2, 4, 4, 2, 8), ncol = 2, byrow=TRUE)

par(mar = c(4, 4, 1, 1)) # Reduce empty margins
plot(ref, xlab=expression(f[1]), ylab=expression(f[2]),
  panel.first=grid(nx=NULL), pch=23, bg="gray", cex=1.5)
points(A, pch=1, cex=1.5)
points(B, pch=19, cex=1.5)
legend("topright", legend=c("Reference", "A", "B"), pch=c(23,1,19),
  pt.bg="gray", bg="white", bty = "n", pt.cex=1.5, cex=1.2)
```

Assuming minimization of both objectives, $A$ is better than $B$ in terms of
Pareto optimality.  However, both `igd()` and `avg_hausdorff_dist()`
incorrectly measure $B$ as better than $A$, whereas `igd_plus()` and
`hypervolume()` correctly measure $A$ as better than $B$ (remember that
hypervolume must be maximized) and `epsilon_additive()` measures both as
equally good (epsilon is weakly Pareto compliant).

```{r, results='asis'}
knitr::kable(
  data.frame(
    A = c(
        igd(A, ref),
        avg_hausdorff_dist(A, ref),
        igd_plus(A, ref),
        epsilon_additive(A, ref),
        hypervolume(A, ref=10)),
    B = c(
        igd(B, ref),
        avg_hausdorff_dist(B, ref),
        igd_plus(B, ref),
        epsilon_additive(B, ref),
        hypervolume(B, ref=10)),
    row.names=c("IGD", "Hausdorff", "IGD+", "eps+", "HV")), format = "html") %>%
  kableExtra::kable_styling(position = "center",
    bootstrap_options = c("striped", "condensed", "responsive"))
```

# References
